
```{r, include = FALSE}
ottrpal::set_knitr_image_path()
```

# Introduction

With the increase in availability of AI chatbots, it is a brave new world in software development. No one is entirely certain what the future will be for those of us who work with code. There are concerns about AI bots replacing programmers or cutting their workload dramatically, resulting in a loss of employment opportunities. There are also serious data privacy concerns as well as other ethical concerns around training data sets. These worries can't be brushed aside lightly. However, this is also a moment of great excitement and discovery as widespread usage of AI bots changes the foundations of computers, data science, and how we interact with information. We think there is much to be optimistic about as we enter this new era of information science.

We do not believe AI chatbots will replace software developers. An AI is only as good as the training data behind it and the human using it. We _do_ think the roles programmers play in software development are changing, and some of the skills needed to work in this new world are slightly different from the ones programmers used to need. Now instead of needing to know programming syntax (or how to find the proper syntax using online communities), a programmer instead needs to understand how to craft a query in order to direct an AI bot to build the workflow they need. Other skills - like how to troubleshoot and debug code, or how to document your code so other can use it - will still be valuable and can be enhanced by the use of AI. AI also creates new niches and opportunities for tasks that were tedious, time-consuming, or not possible before. It is a really exciting time to work in programming.

We used AI bots to develop and write much of this course (although, ironically, not the introduction you just read). We firmly believe this new technology can be used in ethical ways and we put into practice all the rules and suggestions offered over the following modules. It is up to the human using it to make sure that AI is used properly and transparently.


::: {.reflection}

We asked OpenAI's ChatGPT whether AI chatbots would replace human software developers and programmers, and this was the response:

"No, ChatGPT is not replacing human software developers and programmers. While ChatGPT is capable of generating code, it is not a substitute for the skills and expertise of human programmers.

Software development is a complex process that requires a deep understanding of programming languages, algorithms, and software design principles. Human programmers have the ability to not only write code, but also to understand the nuances of different programming languages and frameworks, debug code, and optimize performance.

Furthermore, programming is not just about writing code - it also involves understanding and translating business requirements into technical solutions, collaborating with other team members, and maintaining code over time.

While ChatGPT can be used to automate certain aspects of software development, it cannot replace the creativity, problem-solving skills, and domain expertise of human software developers and programmers."
:::

```{r}
devtools::session_info()
```
