
```{r, include = FALSE}
ottrpal::set_knitr_image_path()
```

# Introduction

The increasing availability of AI chatbots is creating a new landscape in software development. While the future of those who work with code is uncertain, there are valid concerns about the potential for AI bots to replace or significantly reduce the workload of programmers, resulting in a loss of job opportunities. Additionally, there are significant ethical and privacy concerns regarding the use of training data sets to teach these AI systems.

Despite these worries, the widespread adoption of AI bots is ushering in a new era of computing, data science, and information interaction that presents many exciting possibilities. As AI bots become more ubiquitous, they have the potential to change the way we work with information and interact with machines. The rapid development of AI technology is creating exciting opportunities for programmers to develop innovative solutions to previously intractable problems.

Ultimately, while the future of software development is unclear, there is much to be optimistic about in this brave new world. The advancements in AI and chatbot technology present an opportunity for programmers to work alongside algorithms to create innovative solutions that were previously unimaginable. As with any technological revolution, there will be challenges, but it is up to us to work together to ensure that we use these tools ethically and responsibly, and to create a future that benefits everyone.

We do not believe AI chatbots will replace software developers. An AI is only as good as the training data behind it and the human using it. We _do_ think the roles programmers play in software development are changing, and some of the skills needed to work in this new world are slightly different from the ones programmers needed before. Rather than being solely focused on programming syntax, programmers may need to develop a more refined ability to construct queries that enable AI bots to create the desired workflows. Other critical skills, such as debugging, troubleshooting, and documentation, will remain valuable, and can be enhanced through the use of AI. Moreover, AI creates new opportunities and niches for tasks that were previously difficult, time-consuming, or impossible. As a result, this is an exciting time for those working in programming, as they have the potential to explore new and innovative solutions. 

We used AI bots to develop and write much of this course. We firmly believe this new technology can be used in ethical ways and we put into practice all the rules and suggestions offered over the following modules. It is up to the human using it to make sure that AI is used properly and transparently.


::: {.reflection}

We asked OpenAI's ChatGPT whether AI chatbots would replace human software developers and programmers, and this was the response:

"No, ChatGPT is not replacing human software developers and programmers. While ChatGPT is capable of generating code, it is not a substitute for the skills and expertise of human programmers.

Software development is a complex process that requires a deep understanding of programming languages, algorithms, and software design principles. Human programmers have the ability to not only write code, but also to understand the nuances of different programming languages and frameworks, debug code, and optimize performance.

Furthermore, programming is not just about writing code - it also involves understanding and translating business requirements into technical solutions, collaborating with other team members, and maintaining code over time.

While ChatGPT can be used to automate certain aspects of software development, it cannot replace the creativity, problem-solving skills, and domain expertise of human software developers and programmers."
:::

```{r}
devtools::session_info()
```
