

```{r, include = FALSE}
ottrpal::set_knitr_image_path()
```

# Writing Code

In today's digitally-driven world, coding has become an essential skill. It is used to create websites, mobile applications, and sophisticated software systems. Contrary to popular belief, coding is not only for those who possess an innate talent for programming. With the right resources and a willingness to learn, anyone can become proficient in coding.

Practice is crucial to mastering the art of coding, and those who are willing to commit themselves to the learning process can achieve great results. Writing code can also be a rewarding and fulfilling hobby, too. Creating your own software, building a website, or automating a task can be immensely satisfying, and can provide a sense of accomplishment and pride.

Moreover, writing code can be a form of creative expression. Just as musicians use notes to compose music and artists use paint to create visual art, programmers use code to create digital experiences. From building websites to creating video games, code can be used to bring ideas to life in exciting and innovative ways.

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "The dinosaur cartoon is wearing a beret and holding a paintbrush that is is using to paint code onto an easel."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1MCNeSO4aOm1iESWDLOGTcx3aLEbnu8UttV0QGVAeafE/edit#slide=id.g230c4a00725_1_0")
```



Ultimately, the ability to write code is a valuable asset in today's world, with increasing demand for skilled programmers in a wide range of industries. Whether as a career or hobby, coding can be both personally and professionally rewarding. The accessibility and versatility of coding make it an excellent skill for anyone to learn, regardless of their background or previous experience. 


## Writing Code with AI

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "The dinosaur cartoon says in a speech bubble, 'With AI, I am unstoppable!'. The dinosaur is a T-rex and is now holding clamps that extend its arms."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1MCNeSO4aOm1iESWDLOGTcx3aLEbnu8UttV0QGVAeafE/edit#slide=id.g22e0106807c_0_83")
```


AI-powered programming tools and technologies are revolutionizing the way we approach software development. 

1. _AI can generate code snippets._ An AI-powered code editor can suggest more efficient algorithms or highlight potential errors in real-time. It can also help developers to reproduce bugs more easily and to test potential fixes more quickly. This can save developers a significant amount of time and effort, and it can help them to debug their code more effectively.

1. _AI can make programming more accessible._ With AI-generated code and automated programming tools, individuals who are not experienced developers can still create software solutions. This has the potential to democratize programming and lead to new innovations from a wider range of people.

1. _AI is changing the skills required for writing code._ Rather than simply memorizing syntax, developers need to focus on developing their creativity, intuition, and problem-solving skills. Additionally, programmers must also hone their abilities in architecture design and project management, as these skills are becoming increasingly important in the modern software development landscape.

As AI continues to evolve, it is likely to have an even greater impact on the way we develop software in the future.

:::{.ethics}
**As a programmer using AI, it is up to you to be responsible for what you create.** You must rigorously test any code you write. It is your job to make sure any code you create with AI is not malicious and works as expected.

You must also take care that you are only using code that you're authorized to use. This means asking the AI chatbot where code snippets are from and tracking down those citations and authorizations yourself (instead of just trusting the AI that everything is okay).
:::

## Tips for Coding with AI

Using AI to write code is as much of an art as writing your own code.  Following a few rules and suggestions will make your experience much more enjoyable and productive.

**Remember that AI cannot read your mind**. You need to tell it what you're thinking, and it may take several rounds of question-and-answer before you get something you can work with. 

* _Be clear about your goals._ What do you want the AI to do? Be as specific as possible.

* _Start with a simple task._ Don't try to do too much too soon. Start with a simple task and then gradually work your way up to more complex tasks. You wouldn't write a complete program all at once, and you shouldn't have AI do it, either.

* _Provide enough data._ The more data (and detail) you can give the AI, the better it will be able to learn and perform the task you want it to do.

* _Be patient._ Querying an AI can take time. Don't expect it to be perfect right away.

**Test everything AI writes**. You are responsible for the code that AI generates, even if you didn't write it yourself. AI can be biased, just like humans, and this could lead to sub optimal development. There is a concern that AI may focus development on areas that optimize certain metrics but not overall quality, effectiveness, or user needs. Human judgment is still needed to ensure that AI is used in a responsible and ethical way.

* _Test your results._ Make sure the AI is actually doing what you want it to do. AI systems are trained on data, and if the data is not representative of the real world, the AI system may not be able to generate code that works correctly. It is important to test the code to make sure that it is working as intended and that it is producing the desired results. Inadequate testing could lead to low quality, buggy code. Since AI generated code would be highly automated, it requires strong testing procedures and safety oversight to ensure it meets standards before deployment. 

* _Make sure your code is secure._ AI systems can be used to generate code that is vulnerable to attack, and it is important to test the code to make sure that it is secure. This can help to prevent security breaches from occurring. 

* _Use a variety of AI models._ There are many different AI models available. Experiment with different models to see which one works best for your task.

* _Get feedback from others._ Ask other people to review your code and give you feedback. This can help you identify and fix any problems with your code.

## Example One: Creating a Function



## Example Two: Figuring Out a Common Task

Sometimes you have a task that needs done that is fairly routine. You might not have done it before, but plenty of people in the world have. In this case, you don't want to reinvent the wheel, but you may not know exactly where to start. AI is a great option for figuring out the basic workflow steps.

A very common task for anyone working with genomic samples is to combine multiple files in the vcf format. Perhaps you've been asked to do this and then create a separate file that contains the names of the samples. You turn to [ChatGPT](https://chat.openai.com/) for help.

:::{.query}
"Can you write a program that merges vcf files and then outputs the sample names?"
:::

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "ChatGPT generates code that allows you to merge vcf files and output the sample names."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1MCNeSO4aOm1iESWDLOGTcx3aLEbnu8UttV0QGVAeafE/edit#slide=id.g22e0106807c_0_192")
```




Great! Of course, maybe you're not comfortable using Python, or you try this and can't get it to work. You can take advantage of the "Regenerate Response" option in ChatGPT to get another solution.

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "You can use the 'Regenerate Response' button on ChatGPT to find a different way to accomplish your task."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1MCNeSO4aOm1iESWDLOGTcx3aLEbnu8UttV0QGVAeafE/edit#slide=id.g230c4a00725_31_30")
```



:::{.query}
"What's another way I can do this?
:::


```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "ChatGPT generates another way to do the same task."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1MCNeSO4aOm1iESWDLOGTcx3aLEbnu8UttV0QGVAeafE/edit#slide=id.g230c4a00725_31_34")
```



After you find a solution that works for you, you should always make a habit of asking the AI where it got the code. Never, ever, ever run code without knowing where it came from.

:::{.query}
"Did you use any libraries or repos in this code?"
:::

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "ChatGPT responds with the library and basic Unix tools needed to run the generated code."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1MCNeSO4aOm1iESWDLOGTcx3aLEbnu8UttV0QGVAeafE/edit#slide=id.g230c4a00725_31_38")
```




You can also specify some of the features you'd like in your code, or where you might want the code to come from. Maybe you want to practice using GitHub repositories, so you specifically request it.

:::{.query}
"Can you write a program that merges vcf files and then outputs the sample names using a github repo?"
:::

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "ChatGPT generates a script using the vcflib repo from GitHub."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1MCNeSO4aOm1iESWDLOGTcx3aLEbnu8UttV0QGVAeafE/edit#slide=id.g230c4a00725_31_42")
```




If you know you've used something like a GitHub repository, you can ask ChatGPT for the citation. Just be aware that you **need** to verify that any citation it gives you should be checked independently. Unfortunately, AI bots can still respond with imaginary journal articles. (Assume that if you can't find it, it doesn't exist.)

:::{.query}
"What's the citation for the repo used in this code?"
:::

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "ChatGPT responds with the GitHub repository used in the code, as well as the creators and a possible citation."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1MCNeSO4aOm1iESWDLOGTcx3aLEbnu8UttV0QGVAeafE/edit#slide=id.g230c4a00725_31_46")
```





## Example Three: Planning Your Program

Let's say we want to create software that plays the card game blackjack. 
You can jump right in and ask your AI chatbot to write the code, but you're unlikely to be successful and _incredibly_ likely to get frustrated. Before you start interacting with the AI, you need to plan out what you want to do.

Everyone needs a plan of attack when building software. You should start by identifying what your software will do, then you can figure out the steps necessary to make that happen. 

We'll ask [Claude-instant](https://poe.com/Claude-instant) (as implemented by Poe) how we can go about creating the necessary code for our game.

:::{.query}
"How would I go about writing a program that can play blackjack?"
:::

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "Claude-instant laid out seven steps that can be used to build a simple blackjack program. It also offered several ideas for additional features."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1MCNeSO4aOm1iESWDLOGTcx3aLEbnu8UttV0QGVAeafE/edit#slide=id.g230c4a00725_31_24")
```



That seems like a really good roadmap for building a basic game, but if you're ambitious (and of course you are), you might want to give your game a little extra something. Claude-instant has some ideas for that, too.

:::{.query}
"What are some other features I could add to this game?"
:::

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "Claude-instant offered eleven additional features that could be added to your blackjack game if you wanted to make it more complex, including different types of betting strategies and surrender opportunities."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1MCNeSO4aOm1iESWDLOGTcx3aLEbnu8UttV0QGVAeafE/edit#slide=id.g230c4a00725_31_8")
```



At this point, you are ready to begin writing your program. You have a lot of tasks to do, so you have to pick a logical starting point. If you're uncertain, you can ask AI for advice on your first task.

:::{.query}
"What should be my first coding task for this program?"
:::

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "Claude-instant suggests the first task is to create a data structure that represents a deck of cards."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1MCNeSO4aOm1iESWDLOGTcx3aLEbnu8UttV0QGVAeafE/edit#slide=id.g230c4a00725_31_68")
```



In this case, the AI went beyond just suggesting a first task and also offered several options for how to do it as well. 

## Example Four: Selecting a Language

You will need to choose the language for your software. If you don't, AI will likely pick a language for you, which may or may not be one you want. You might choose a language based on:

* Your preexisting knowledge or experience with the language
* A specific use case where the language is commonly used
* The language(s) your colleagues or coworkers use


You can find a description of other languages you might encounter [here](Appendix.html).

Let's ask Claude-instant again for some recommendations on the appropriate languages for our proposed blackjack game from the above example. (Pro tip: the various AI bots all keep track of your previous queries, so you don't have to restate your goal for the software if you're continuing the conversation.)

:::{.query}
"What would be a good coding language to create this game of blackjack?"
:::

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "Claude-instant suggested seven possible languages for the blackjack program and gave a brief description of their pros and cons."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1MCNeSO4aOm1iESWDLOGTcx3aLEbnu8UttV0QGVAeafE/edit#slide=id.g230c4a00725_31_4")
```



Claude-instant actually gave us more information than we anticipated. If it hadn't stated pros and cons while offering language suggestions, we could ask for them directly. This could be helpful if you're deciding between two options like, say, Python or Java.

:::{.query}
"What are the pros and cons of using Python vs Java for this project?"
:::

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "Claude-instant directly compares Python and Java as languages for the blackjack game, suggesting Python is best for a simple game and Java is better if you want to build a more complex game."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1MCNeSO4aOm1iESWDLOGTcx3aLEbnu8UttV0QGVAeafE/edit#slide=id.g230c4a00725_31_0")
```




## Limitations

All of the AI chatbots we use are still in the process of being developed



## Hands On Exercise 



```{r}
devtools::session_info()
```


