

```{r, include = FALSE}
ottrpal::set_knitr_image_path()
```

# Writing Code

In today's digitally-driven world, coding has become an essential skill. It is used to create websites, mobile applications, and sophisticated software systems. Contrary to popular belief, coding is not only for those who possess an innate talent for programming. With the right resources and a willingness to learn, anyone can become proficient in coding.

Practice is crucial to mastering the art of coding, and those who are willing to commit themselves to the learning process can achieve great results. Writing code can also be a rewarding and fulfilling hobby, too. Creating your own software, building a website, or automating a task can be immensely satisfying, and can provide a sense of accomplishment and pride.

Moreover, writing code can be a form of creative expression. Just as musicians use notes to compose music and artists use paint to create visual art, programmers use code to create digital experiences. From building websites to creating video games, code can be used to bring ideas to life in exciting and innovative ways.

[FIGURE SHOWING CODING AS A CREATIVE ART???]

Ultimately, the ability to write code is a valuable asset in today's world, with increasing demand for skilled programmers in a wide range of industries. Whether as a career or hobby, coding can be both personally and professionally rewarding. The accessibility and versatility of coding make it an excellent skill for anyone to learn, regardless of their background or previous experience. 


## Writing Code with AI

```{r, fig.align='center', out.width="100%", echo = FALSE, fig.alt= "The dinosaur cartoon says in a speech bubble, 'With AI, I am unstoppable!'. The dinosaur is a T-rex and is now holding clamps that extend its arms."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1MCNeSO4aOm1iESWDLOGTcx3aLEbnu8UttV0QGVAeafE/edit#slide=id.g22e0106807c_0_83")
```

AI-powered programming tools and technologies are revolutionizing the way we approach software development. 

1. _AI can generate code snippets._ An AI-powered code editor can suggest more efficient algorithms or highlight potential errors in real-time. It can also help developers to reproduce bugs more easily and to test potential fixes more quickly. This can save developers a significant amount of time and effort, and it can help them to debug their code more effectively.

1. _AI can make programming more accessible._ With AI-generated code and automated programming tools, individuals who are not experienced developers can still create software solutions. This has the potential to democratize programming and lead to new innovations from a wider range of people.

1. _AI is changing the skills required for writing code._ Rather than simply memorizing syntax, developers need to focus on developing their creativity, intuition, and problem-solving skills. Additionally, programmers must also hone their abilities in architecture design and project management, as these skills are becoming increasingly important in the modern software development landscape.

As AI continues to evolve, it is likely to have an even greater impact on the way we develop software in the future.

:::{.ethics}
**As a programmer using AI, it is up to you to be responsible for what you create.** You must rigorously test any code you write. It is your job to make sure any code you create with AI is not malicious and works as expected.

You must also take care that you are only using code that you're authorized to use. This means asking the AI chatbot where code snippets are from and tracking down those citations and authorizations yourself (instead of just trusting the AI that everything is okay).
:::


## Step One: Planning Your Program

You can jump right in and ask your AI chatbot to write the code, but you're unlikely to be successful and _incredibly_ likely to get frustrated. Before you start interacting with the AI, you need to plan out what you want to do.

### Example One: Defining Your Approach

Everyone needs a plan of attack when building software. You should start by identifying what your software will do, then you can figure out the steps necessary to make that happen. 

Let's say we want to create software that 

### Example Two: Selecting a Language

You also need to choose the language for your software. You might choose a language based on:

* Your preexisting knowledge or experience with the language
* A specific use case where the language is commonly used
* The language(s) your colleagues or coworkers use

AI chatbots are familiar with a huge variety of coding languages, which allows them to be used for a variety of tasks. As AI chatbots continue to develop, they are likely to become even more familiar with a wider variety of coding languages. 

Let's ask [Claude-instant](https://poe.com/Claude-instant) (as implemented by Poe) for some recommendations on the appropriate languages for our proposed software from Example One. (Pro tip: the various AI bots all keep track of your previous queries, so you don't have to restate your goal for the software if you're continuing the conversation.)

:::{.query}
"What would be a good coding language...?"
:::


Here are commonly used languages and their typical use cases.

**Python**

Known for its readability, simple syntax, and comprehensive add-on modules. It is often recommended as a beginner-friendly language. Python is used for scientific computing, data analysis, machine learning, and web development. Many Data Scientist roles rely on Python.

**Ruby**

Similar to Python, Ruby is also known for its readability and user-friendly syntax. It is typically used for web development, building dynamic websites, and creating web applications.

**JavaScript**

A popular language for web development, known for its versatility and ease of use. Many of the interactive features of websites rely on JavaScript.

**Java**

Java is widely used for enterprise applications, android mobile apps, and web apps.

**PHP**

PHP is easy to learn but can become complex as the application grows. It is typically used for building dynamic web pages, web applications, and content management systems.

**R**

Used for statistical computing, data analysis, and data visualization, R is a popular choice for researchers and data scientists. Some beginners find its syntax complex, especially due to differences in the two major dialects: "base" R versus the "tidyverse".

**Kotlin**

Kotlin is a relatively new language that is designed to be more concise and expressive than Java. However, its advanced features and concepts may be challenging for beginners. It is used for developing Android mobile apps, web development, and server-side development.

**Swift**

Used for iOS and macOS app development, Swift is known for its strict syntax and steep learning curve.

**C#**

Often used for Windows desktop applications, web applications, and video games, C# can be challenging due to its more complex syntax and strict type system.

**C++**

A powerful language often used for developing high-performance applications and operating systems, C++ can be challenging to learn due to its complexity and the need to manage memory manually.

You can find a description of other languages you might encounter [here](Appendix.html).

## Step Two: Writing And Debugging-As-You-Go

Using AI to write code is as much of an art as writing your own code.  Following a few rules and suggestions will make your experience much more enjoyable and productive.

**Remember that AI cannot read your mind**. You need to tell it what you're thinking, and it may take several rounds of question-and-answer before you get something you can work with. 

* _Be clear about your goals._ What do you want the AI to do? Be as specific as possible.

* _Start with a simple task._ Don't try to do too much too soon. Start with a simple task and then gradually work your way up to more complex tasks. You wouldn't write a complete program all at once, and you shouldn't have AI do it, either.

* _Provide enough data._ The more data (and detail) you can give the AI, the better it will be able to learn and perform the task you want it to do.

* _Be patient._ Querying an AI can take time. Don't expect it to be perfect right away.

**Test everything AI writes**. You are responsible for the code that AI generates, even if you didn't write it yourself. AI can be biased, just like humans, and this could lead to sub optimal development. There is a concern that AI may focus development on areas that optimize certain metrics but not overall quality, effectiveness, or user needs. Human judgment is still needed to ensure that AI is used in a responsible and ethical way.

* _Test your results._ Make sure the AI is actually doing what you want it to do. AI systems are trained on data, and if the data is not representative of the real world, the AI system may not be able to generate code that works correctly. It is important to test the code to make sure that it is working as intended and that it is producing the desired results. Inadequate testing could lead to low quality, buggy code. Since AI generated code would be highly automated, it requires strong testing procedures and safety oversight to ensure it meets standards before deployment. 

* _Make sure your code is secure._ AI systems can be used to generate code that is vulnerable to attack, and it is important to test the code to make sure that it is secure. This can help to prevent security breaches from occurring. 

* _Use a variety of AI models._ There are many different AI models available. Experiment with different models to see which one works best for your task.

* _Get feedback from others._ Ask other people to review your code and give you feedback. This can help you identify and fix any problems with your code.

### Example Three: Chunking It Out

### Example Four: Slowly Adding Complexity

### Example Five: Debugging


## Hands On Exercise 



```{r}
devtools::session_info()
```


