

```{r, include = FALSE}
ottrpal::set_knitr_image_path()
```

# Writing Code: Hands-On Exercise 

Now it’s your turn to try!

## Code

Here's the code you need for question #2.

```
def calculate_average(numbers):
    total = 0
    for number in numbers:
        total += number
    average = total / len(numbers)
    return average

numbers = [1, 2, 3, 4, "5"]
average = calculate_average(numbers)
print("The average of the numbers is:", average)
```

## Questions

1. Write a prompt to create a function that replaces all instances of the string "three" with the number 3, then create a prompt that allows you to modify the function so that "three", "Three", and "THREE" are all replaced with the number 3.

1. Write a query that helps you debug the code snippet from above.

1. Create a query that explains how to retrieve data from an online database. Don't forget to find out what the source of the code is.

1. Write a prompt to plan out a program that creates and plays a game of Rock, Paper, Scissors. You should create a multi-step plan as well as figure out what your first coding task should be.

1. Create a prompt that identifies the best coding language for the above game.



```{r}
devtools::session_info()
```


